Your Main Manuscript Title

### 2018-08-21 papers/paper/paper-supplemental.tex
Add paper-suplemental.tex with more deptn on OB and BB derivations

\errorstopmode

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Masters/Doctoral Thesis 
% LaTeX Template
% Version 2.4 (22/11/16)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Version 2.x major modifications by:
% Vel (vel@latextemplates.com)
%
% This template is based on a template by:
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Template license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
11pt, % The default document font size, options: 10pt, 11pt, 12pt
%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
english, % ngerman for German
singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
%toctotoc, % Uncomment to add the main table of contents to the table of contents
%parskip, % Uncomment to add space between paragraphs
%nohyperref, % Uncomment to not load the hyperref package
headsepline, % Uncomment to get a line under the header
chapterinoneline, % Uncomment to place the chapter title next to the number on one line
%consistentlayout, % Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
]{MastersDoctoralThesis} % The class file specifying the document structure

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters

\usepackage{lmodern} % Use the Palatino font by default

\usepackage[backend=bibtex,style=numeric,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{thesis-articles.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography

% Elliott-added packages
\usepackage{amsmath} % for align and other environments
\usepackage{listings}
\usepackage{float}   % for H figure specifier
\usepackage{indentfirst} % to indent after sections

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper, % Change to letterpaper for US letter
	inner=2.5cm, % Inner margin
	outer=2.5cm, % Outer margin
	bindingoffset=.0cm, % Binding offset
	top=1.5cm, % Top margin
	bottom=1.5cm, % Bottom margin
	%showframe, % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

\thesistitle{Simulating dynein'’s powerstroke using Brownian dynamics} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Dr. David \textsc{Roundy}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{Bachelor of Science} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{Elliott \textsc{Capek}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{Physics} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{http://www.university.com}{Oregon State University}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\department{\href{http://physics.oregonstate.edu}{Oregon State University Department of Physics}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\group{\href{http://physics.oregonstate.edu/~roundyd/}{Roundy Research Group}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\faculty{} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

@@

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
\begin{center}

\vspace*{.06\textheight}
{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
\textsc{\Large Undergraduate Thesis}\\[0.5cm] % Thesis type

\HRule \\[0.4cm] % Horizontal line
{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
\HRule \\[1.5cm] % Horizontal line

\begin{minipage}[t]{0.4\textwidth}
\begin{flushleft} \large
\emph{Author:}\\
{\authorname} % Author name - remove the \href bracket to remove the link
\end{flushleft}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{flushright} \large
\emph{Supervisor:} \\
{\supname} % Supervisor name - remove the \href bracket to remove the link  
\end{flushright}
\end{minipage}\\[3cm]

\vfill

\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
\textit{in the}\\[0.4cm]
\groupname\\\deptname\\[2cm] % Research group name and department name

\vfill

{\large \today}\\[4cm] % Date
%\includegraphics{Logo} % University/department logo - uncomment to place it

\vfill
\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}
\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
Dynein is a motor protein which transports cargo along tracks inside the cell. Like related motor proteins kinesin and myosin, dynein uses cellular energy to take steps with its two foot domains. Unlike kinesin or myosin, dynein's stepping pattern is highly varied: it can take steps between zero and 60nm in both the forwards and backwards directions. It is believed that dynein takes such broad, stochastic steps because its large size and several elastic regions make it more influenced by Brownian motion. To test this, we model the motor as a 2D system of elastic hinges, then simulate this model using Brownian dynamics. Preliminary results indicate such a model is capable of taking steps between zero and 25nm. These results give hope that, with further tweaking, the model may be able to generate both larger steps and backwards steps.
\end{abstract}
\newpage

%% %----------------------------------------------------------------------------------------
%% %	ACKNOWLEDGEMENTS
%% %----------------------------------------------------------------------------------------

%% \begin{acknowledgements}
%% \addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
%% The acknowledgments and the people to thank go here, don't forget to include your project advisor\ldots
%% \end{acknowledgements}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents
\newpage

\listoffigures % Prints the list of figures
\newpage

%% \listoftables % Prints the list of tables
\newpage

%% %----------------------------------------------------------------------------------------
%% %	ABBREVIATIONS
%% %----------------------------------------------------------------------------------------

%% \begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)

%% \textbf{MT} & \textbf{M}icro\textbf{t}ubule\\
%% \textbf{MTBD} & \textbf{M}icro\textbf{t}ubule \textbf{B}inding \textbf{D}omain\\

%% \end{abbreviations}

%% %----------------------------------------------------------------------------------------
%% %	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%% %----------------------------------------------------------------------------------------

%% \begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table

%% % The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it

%% Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
%% %Constant Name & $Symbol$ & $Constant Value$ with units\\

%% \end{constants}

%% %----------------------------------------------------------------------------------------
%% %	SYMBOLS
%% %----------------------------------------------------------------------------------------

%% \begin{symbols}{lll} % Include a list of Symbols (a three column table)

%% $a$ & distance & \si{\meter} \\
%% $P$ & power & \si{\watt} (\si{\joule\per\second}) \\
%% %Symbol & Name & Unit \\

%% \addlinespace % Gap to separate the Roman symbols from the Greek

%% $\omega$ & angular frequency & \si{\radian} \\

%% \end{symbols}

%% %----------------------------------------------------------------------------------------
%% %	DEDICATION
%% %----------------------------------------------------------------------------------------

%% \dedicatory{For/Dedicated to/To my\ldots}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------
\chapter{Introduction}
\label{chap:Introduction}

\blockquote{An embryo in its spherical blastula stage has spherical symmetry [...]. But a system which has spherical symmetry, and whose state is changing because of chemical reactions and diffusion, will remain spherically symmetrical for ever [...]. It certainly cannot result in an organism such as a horse, which is not spherically symmetrical. \textbf{Alan Turing, The Chemical Basis of Morphogenesis, 1952}}

Cells need to create asymmetric, heterogeneous structures inside themselves in order to function. However, diffusive processes cannot create this asymmetry. How then do cells create the heterogeneous structures they need to exist? Motor proteins are part of the answer. Motor proteins are structures which use energy to create directed motion in cells. In broad terms, they are mechanical creators of asymmetry. This thesis explores a hypothesis about how one such motor protein, dynein, generates directed motion.\\

\section{Motivation}
Motor proteins are cellular structures which use chemical energy to do work and generate motion. Kinesin, myosin and dynein are the three main motor proteins in cells. Though these proteins have significantly different ancestors \cite{hook-family}, they all evolved into a single similar structure. In biology when a structure is evolved three separate times, it is a sign that nature has found an efficient way of solving a problem. How, then, do motor proteins work?\\

The basic process is simple. Kinesin, myosin and dynein each use two binding domains to attach to a surface. These binding domains are connected via a linker domain. Each motor protein is an engine which undergoes cycles, where each cycle moves the motor across a surface. In a single cycle a binding domain unbinds from a surface, tenses its foot to kick forward, rebinds, then untenses its foot to pull the whole motor forward. This pattern is repeated by both feet to generate motion \cite{kinesin-cycle, myosin-cycle}.\\

Although the three motor protein families operate using the same stepping cycle, their behaviors differ significantly. Kinesin and myosin are predictable motors which step in the forward direction with a regular step size. In contrast, dynein's stepping pattern varies in both size and direction. The motor can take 64nm steps forward or 32nm steps backward, and often steps sideways as well \cite{weihongpaper}. How can motors which implement the same release, lever, bind, lurch stepping mechanism behave so differently?\\

One possibility is that dynein's size and organization are responsible for the motor's odd stepping pattern. Dynein is much larger than kinesin or myosin, so its feet may need to move significantly further during each cycle. Due to the random nature of Brownian motion from water particle collisions, there may be significantly more variance in the path taken by dynein's feet compared to other motor feet. This could explain the large variance in step size. Another contributing factor could be that dynein is so large that it can twist and contort during its step, allowing varying and even backwards steps, which a small motor would be too stiff to achieve.\\

This thesis describes an experiment which aims to test whether Brownian motion and flexibility cause dynein's unique stepping pattern. To do this, we construct a mathematical model of the dynein motor which exhibits flexibility and feels Brownian forces. We then simulate the model and compare its behavior to real dynein.\\

%% Dynein:
%% feet which bind to track
%% motor connected to feet which generates the force
%% -----------
%% ATP binding causes dissociation
%% Also causes forward kick
%% ATP hydrolysis
%% rebinding causes ADP rejection and forward lurch
%% restart

%% Kinesin:
%% feet which bind to track
%% connected via lever arm to cargo binding domain
%% -----------
%% ATP binding causes forward kick
%% ATP hydrolysis
%% MT binding, ADP, Pi rejection
%% Other foot unbinds

%% Myosin:
%% Feet which bind to track
%% connected via lever arm to cargo binding domain
%% ----------
%% actin binding causes Pi rejection, strengthening binding
%% binding causes foot to change relative to lever, shifting the motor forward
%% ATP binding causes dissocaition and undoing the powerstroke

\section{What is dynein?}
\subsection{Basics}
Dynein is a cellular structure which converts chemical energy to mechanical energy. It does so by reacting with adenosine triphosphate (ATP) to take steps along long cellular tracks known as microtubules (MTs). Dynein's mechanical energy is used for various cellular functions, including to transport molecules around the cell, pull chromosomes during division, and move cellular propellers known as cilia \cite{cianfroccoreview}.\\

Dynein is a protein with many smaller domains. A depiction of the protein and its labeled subdomains is shown in Figure (\ref{dynein-artist-rendition}). Each motor is a homodimer made up of two identical monomer proteins. Each monomer has a binding domain which binds to the microtubule, a motor domain which binds ATP, a tail domain which connects to the other monomer, a stalk which connects the binding and motor domains, and a linker which connects the motor and tail domains. The tail domain is also responsible for binding to cargo.\\

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/dynein-artist-rendition.jpg}
  \caption{Artist's rendition of dynein bound to a microtubule.}{Two heavy chains (cyan and blue) and several tail proteins (pink) come together to form the dynein motor complex. One monomer's binding domain is bound to the microtubule (green), and the other is diffusing. ATP hydrolysis sites are labeled glowing green. Figure modified. Source: Lander Lab, The Scripps Research Institute \cite{landerartistsrendition}.}
  \label{dynein-artist-rendition}
\end{figure}

Dynein is believed to move by using the energy of ATP to cycle through a sequence of states, each with different spatial relations to the microtubule. By cycling forward through this sequence, the molecule will end up in the same state it began, but moved forward a small amount. This sequence of states is known as the mechanochemical cycle \cite{cianfroccoreview}. The goal of this project is to create a mathematical model of the mechanochemical cycle and verify that it can reproduce experimental dynein stepping data. There are several types of dynein which all behave slightly differently. This project will focus on cytoplasmic dynein-1, hereby referred to as dynein.\\

\subsection{Why do cells need motor transport?}
Cells are organized, heterogenous structures which respond quickly to their environments. This means that cells require a mechanism for rapidly moving components to precise locations within the cell. This can be a challenge, since cells are fairly large compared to the proteins which compose them. A human fibroblast cell has a volume of roughly 2000 $\mu\text{m}^3$\cite{fibroblastvolume}, corresponding to roughly $8 \mu \text{m}$ in diameter. In comparison, human hemoglobin has a diameter of roughly $5 \text{nm}$ (PDB id 5ME2), a $10^3$ factor decrease in size.\\

Diffusion, or random motion of molecules due to collisions with solvent (i.e. Brownian motion), is one possible process cells could use to transport biomolecules. Diffusion has two problems: it is slow and nondirected. The expected distance covered for one-dimensional Brownian motion after time t was found by Einstein to be \cite{einstein-paper}:

\begin{equation}
  \label{diffusion-equation}
  \langle x\rangle = \sqrt{2Dt}
\end{equation}
%
where D is the diffusion constant describing the diffusability of the biomolecule. For a medium-sized (140kD, 1D = weight of H atom) protein with a diffusion constant $D = .2 \mu \text{m}^2/\text{s}$ \cite{diffusionconstants}, it would take about a month to travel across a millimeter-sized cell. In contrast, it would take dynein, which travels at roughly 100nm/s \cite{weihongpaper}, a much more reasonable three hours to do the same. \\%% Figure (\ref{fig:diffusion_vs_dynein}) demonstrates how motor transport quickly outstrips diffusion in a cell.\\

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/diffusion_vs_dynein.png}
%%   \caption{Passive diffusion vs motor transport rate}{Plot comparing the diffusion rate of a 140kD protein with a diffusion constant $D = .2 \mu m^2/s$ with a dynein motor travelling at 100nm/s.}
%%   \label{fig:diffusion_vs_dynein}
%% \end{figure}

Another advantage of motor transport is that microtubules can be highly specific in their location in the cell. During cell division chromosomes line up at the center of each cell in a plane. Such a geometrically precise shape requires some sort of directed motion which a random process could never provide.\\

\subsection{What role does dynein play in the cell?}
One of dynein's primary jobs is to transport cellular cargos such as organelles, vesicles, mRNA, cytoskeletal filaments and certain proteins from place to place in the cell. Dynein also plays a role in positioning and breaking down the nucleus, cell death, spindle formation and the placement of other important cellular structures \cite{valetoolbox}.\\

Dynein walks along cellular tracks called microtubules (MTs). MTs are long polymers of alternating $\alpha-$ and $\beta-$ tubulin subunits. MTs are polar, meaning they have a distinct directionality based on the orientation of the tubulin proteins which comprise them. One pole, called the minus end, is where they initiate formation, typically at a MTOC (microtubule organizing center) around the center of the cell. The other pole, the plus end, is where they grow and shrink from. Kinesin, another type of motor protein, typically walks towards the plus end of the MT. Dynein is unique in that it walks towards the minus end, making it a minus end-directed motor. This lends it to a particularly important function inside neurons known as retrograde transport.\\

Neurons have cell bodies at their centers and axons which grow outwards. Axons are long, narrow structures extending up to a meter in length in humans. Cell bodies contain nuclei and important organelles for synthesizing proteins, but growth of axon tips is vital for neuron function. This means bidirectional motor transport between axon tips and cell bodies is very important in neurons, since diffusion would not be quick enough. Retrograde transport is transport from the axon tip to the cell body. Cargo includes vesicles full of proteins ready to be broken down in the cell body and microtubule fragments to be returned from the axon tip \cite{neuroanatomy}. Because microtubule minus ends are found in the cell body, dynein's minus end-directed nature makes it the only protein capable of retrograde transport. An interesting question is, if proteins like dynein are synthesized in the cell body but needed at the axon tip to perform retrograde transport, how do dynein get to the tip? The answer is shown in Figure (\ref{retrograde-transport}): kinesin, a plus end-directed motor, takes dynein to the tip from the cell body.\\

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/retrograde_transport.jpg}
  \caption{Dynein transport to neuron growth cones.}{Dynein and kinesin collaborate to move cargo between the neuron body and axon tip. The oppositely-directed motors take turns transporting each other, along with cargo for building the axon growth cone. Image from \cite{valetoolbox}.}
  \label{retrograde-transport}
\end{figure}

\subsection{Dynein structure}
As shown in the crystal structure in Figure (\ref{fig:pdb-motm-dynein}), a dynein motor is a fusion of two identical subunits, or monomers, each with binding, stalk, motor and tail subdomains. Each of these subunits has a unique purpose for the motor.\\

Dynein is a fusion of two identical proteins, referred to as monomers. A single monomer, depicted in (\ref{fig:pdb-motm-dynein}), is composed of a microtubule-binding domain (MTBD), a stalk, a buttress, a motor domain, a linker and a tail. The MTBD is responsible for binding to the microtubule. The stalk links the motor domain to the MTBD and carries information about the state of the motor to the MTBD. The motor domain hydrolyses ATP and performs the main conformation changes involved in the step \cite{FRETstatepaper}. The linker connects the tail to the motor and also changes conformation during the step. The tail is a binding site which interacts with several binding partners, including the tail of a twin monomer.\\

\begin{figure}[h]
  \centering
  \includegraphics[width=.5\textwidth,keepaspectratio]{../figures/pdb-motm-dynein}
  \caption{Structure of a dynein monomer.}{A cryo-electron microscopy-inspired image detailing the subcomponents of a single dynein monomer. The microtubule-binding domain (MTBD) connects to the motor domain via a coiled coil known as the stalk. The six AAA regions provide structure for the motor domain, and AAA2 conducts the ATP hydrolysis. The linker is a flexible region which connects the motor domain to the tail. One of the main roles of the tail is connecting a monomer to its twin to form the dynein complex. Image from \cite{pdb-motm-dynein}.}
  \label{fig:pdb-motm-dynein}
\end{figure}

\subsection{Why is dynein interesting?}
Selective domain labeling indicates dynein's stepping pattern is highly varied \cite{reck2006single, weihongpaper}. These studies show dynein's stepping pattern to be stochastic and weakly forward-directed. As shown in Figure (\ref{fig:weihong-steps}), dynein takes many backwards steps, and also many steps with components perpendicular to the microtubule. This indicates the process dynein uses to generate motion has a stochastic component to it. In contrast, kinesin is known to take very regular 8nm steps in the absence of large forces \cite{kinesin-step-size}. For reference, the diameter of a microtubule is roughly 25nm.\\

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.49\textwidth]{../figures/weihong-step-size}
  \includegraphics[width=0.49\textwidth]{../figures/weihong-off-axis-steps}
    \caption{Experimental dynein stepping behavior histograms.}{Left: Size of steps taken by dynein motor parallel to microtubule. Positive axis corresponds to steps towards the microtubule minus-end. Right: Size of ``off-axis'' steps taken by the dynein motor perpendicular to the microtubule. Figure modified from \cite{weihongpaper}.}
  \label{fig:weihong-steps}
\end{figure}

Kinesin and dynein, although of very different evolutionary origin \cite{hook-family}, have very similar stepping cycles. Both proteins bind bipedally to microtubules. Both proteins use ATP hydrolysis to leverage against a linker domain, pushing the motor domain forward. Both proteins couple their nucleotide and MT-binding states \cite{vale-kinesin-cycle}. How, given the large number of similarities between these motors, does dynein achieve such a radically different stepping pattern than kinesin? Elucidating how the structural differences between the motors give rise to the differences in dynamics would provide lots of important information on how the machines fine-tune their stepping.\\

A major factor which could contribute to differences in stepping is each protein's size. Kinesin proteins are generally 100-200 kD in mass, whereas dynein is on the order of 1mD. As shown in Figure (\ref{fig:kinesin-vs-dynein}), the dynein motor domain is much bigger than the kinesin motor domain. Another key difference is the separation between ATP hydrolysis sites on either protein. Kinesin's ATP hydrolysis site is inside the globular microtubule-binding domain. In contrast, Dynein binds to the MT with its MTBD, colored yellow in Figure (\ref{fig:kinesin-vs-dynein}), and hydrolyses ATP in its motor domain, which is colored cyan. This distance is close to 20nm in length - much longer than kinesin. This large separation might act like a loose ``leash,'' allowing the binding domain to diffuse freely around the microtubule. Kinesin has a much more rigid leash, as shown in the bottom pane of (\ref{fig:kinesin-vs-dynein}). This more rigid leash may prevent diffusion from happening as much.\\

\begin{figure}[H]
  \centering
  \begin{minipage}{0.3\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio]{../figures/kinesin_vs_dynein}
  \end{minipage}
  \begin{minipage}{0.65\textwidth}
    \includegraphics[width=\textwidth,keepaspectratio]{../figures/kinesin_vs_dynein_2}
  \end{minipage}
  \caption{Structural comparison of dynein and kinesin}{Left: Comparison of crystal structures of the motor domains of dynein and kinesin bound to a microtubule. Image from \cite{kinesin-vs-dynein}. Right: Full motor complexes of dynein and kinesin inspired by cryo-EM microscopy. The blue regions on the right image of either protein are the motor domains, which are pictured in the left crystal structures. Not present in the left cyrstal structures are the tail and intermediate chains of either protein. Right image from \cite{kinesin-vs-dynein-2}.}
  \label{fig:kinesin-vs-dynein}
\end{figure}

\section{Dynein Stepping Background}
Biophysicists and molecular biologists have conducted experiments over the past twenty years which shed light on how the dynein motor might walk. The full mechanism is not known, but several known behaviors of dynein can be taken together to construct a model for how the motor walks.\\

\subsection{ATP cycle influences conformational state}
Dynein is known to occupy multiple discrete conformational states depending on its environment. In particular, the concentration of ATP and microtubules is known to alter the shape dynein takes on. A FRET (fluorescence resonance energy transfer) study attached fluorophores to dynein's motor and tail and measured the average distance between the two \cite{FRETstatepaper}. It was found that average distance increased or decreased depending on the nucleotide bound to dynein, indicating the motor takes on different conformations for different bound nucleotides. Dynein unbound to nucleotide, or frozen in ATP or ADP-bound states, had a similar tail-motor distance. However, when the motor was stuck in an ADP-Pi state, with ATP hydrolyzed but phosphate not yet released, the tail-motor distance was increased. This suggests that dynein starts its cycle with no nucleotide, binds ATP, hydrolyzes ATP, enters the ADP-Pi state and changes conformation to a far tail-motor distance, then releases phosphate, moves back to its original close tail-motor conformation, then ejects ADP and starts over.\\

Studies suggest that near/far tail-motor conformation changes are not due to lengthening of the stem linking the two domains, but rather a change in position of the head relative to the tail \cite{FRETstatepaper, burgess-paper}. Cryo-electron microscopy images of the dynein motor shown in Figure (\ref{fig:cryo-em-angles}) demonstrate the differences in conformation in the nucleotide-free and ADP-Pi state. The shifting of the linker domain in the ADP state is known as the powerstroke, and is believed to be the principle cause for dynein's forward step.\\

\begin{figure}[h]
  \centering
  \includegraphics[width=.65\textwidth,keepaspectratio]{../figures/burgess-dynein-angles}\\
  \hspace{1cm}
  \includegraphics[width=.95\textwidth,keepaspectratio]{../figures/linker-swing}
  \caption{Dynein conformation change during a step.}{Top: Cryo-electron microscopy images of axonemal dynein in two different nucleotide states. Top left: dynein frozen in ADP-Pi state. Top middle: dynein without nucleotide. Top right: two states superimposed on one another. Figures from \cite{burgess-paper}. Bottom: X-ray diffraction-inspired image of a motor in two nucleotide states. Bottom left: motor in ADP-Pi state. Bottom right: Motor in ADP state. The linker undergoes an approximately 40 degree swing between these two states known as the powerstroke. Figure from \cite{bhabha-paper}.}
  \label{fig:cryo-em-angles}
\end{figure}

Crystal structures of cytoplasmic dynein-2, a relative of dynein, indicate that ATP-Pi binding to the motor domain induces a conformation change in the motor domain, which alters the angle at which the motor binds to the tail \cite{carter-paper}. Taken together, these results indicate that as the motor binds and hydrolyzes ATP, it undergoes a conformation change which physically moves the motor relative to the tail.\\

%% \textbf{TODO: look at \cite{nicastro} and \cite{schmidt-carter} to see if they can add to this discussion}

\subsection{Microtubule and nucleotide states are coupled}
Dynein is known to bind to the microtubule at a location known as the $\alpha$-$\beta$ polymerization site \cite{redwineMTBDcomplex}. This site occurs roughly every 8nm, indicating dynein is restricted to forward step sizes which are multiples of 8nm. The motor is also capable of stepping sideways \cite{weihongpaper}, indicating total displacements which are not multiples of 8nm.\\

The nucleotide state of the motor is not the only determiner of dynein conformation; microtubule binding state has also been shown to influence conformation. A study on the MTBD-MT complex indicates the stalk connecting the motor and MTBD can take on two major conformations, depending on whether the MTBD is bound to the MT or not \cite{mt-atp-coupling}. The conformation of the stalk has been shown to communicate information between the motor and MTBD. In the presence of microtubules, the ATPase rate of the motor is increased 25-fold. This indicates the stalk communicates MT-binding state of the MTBD to the motor, increasing its ATP hydrolysis rate. The communication happens the other way, as well. Treating dynein with ATP causes MT-binding rate to drop 15-fold, indicating ATP-binding at the motor decreases MT-binding rate at the MTBD.\\

In addition to a bidirectional linking between ATPase rate and MT-binding, stalk state has been shown to directly influence motor conformation \cite{mt-atp-coupling}. When the stalk is allowed to freely transition between bound and unbound states, the motor can also transition between near and far tail/motor states. However, when the stalk is frozen in a particular conformation, the motor loses its transitioning ability. This suggests that the conformations of the two systems are tightly linked.\\

\section{Cianfrocco model for dynein stepping}
\label{sec:cianfrocco-model}
Data presented in the previous sections paints a picture of dynein as a collection of interconnected parts, each cycling through its own collection of states and influencing other parts. Somehow the protein interacts with nucleotides and microtubules in a way that allows it to reliably move through the cell. The Cianfrocco model presented below takes these facts and string them together into cohesive cycles which might explain how dynein walks.\\

Cianfrocco \textit{et. al} \cite{cianfroccoreview} propose dynein achieves motion via transitioning through the cycle of states shown in Fig. (\ref{mech-cycle}). The model claims the primary events in the dynein step are as follows. First, starting with both binding domains attached to the MT, ATP binds to the motor domain, which causes it to change conformation. ATP hydrolysis to ADP-Pi causes further conformation changes. These changes cause the binding domain to release the microtubule, meaning only one binding domain is MT-bound. The free binding domain then undergoes another conformation change known as the ``priming stroke,'' which changes the orientation of the binding domain with respect to the MT. Finally the binding domain rebinds to the MT, causing a final conformation change in the motor which returns it to the beginning of the cycle. The exact order of these events is not precisely known.\\

\begin{figure}[h!]
  \centering
  \includegraphics[width=.90\textwidth,keepaspectratio]{../figures/mechanochemical-cycle.jpeg}
  \newcommand\bothbound{\textcolor{blue}{BOTHBOUND}}
  \newcommand\onebound{\textcolor{red}{ONEBOUND}}
  \newcommand\transition{\textcolor{green}{transition?}}
  %% \begin{tabular}{rll}
  %%   \bothbound &
  %%   1 & \begin{minipage}{5in}
  %%     Bound to microtubule and waiting for ATP to arrive.  The
  %%     conformation is ``post-powerstroke''.
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \bothbound &
  %%   2 & \begin{minipage}{5in}
  %%     ATP has bound to the motor at AAA1, but nothing else has changed.
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \transition &
  %%   3 & \begin{minipage}{5in}
  %%     ``The ring has closed'' in the motor, whatever that means.
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \bothbound &
  %%   4 & \begin{minipage}{5in}
  %%     The affinity to the microtubule is dropped, so it is likely to pop off soon.
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \onebound &
  %%   5 & \begin{minipage}{5in} It has popped off of the microtubule!
  %%     Onebound! The equilibrium angle is also changed.  The ATP has
  %%     been hydrolized, but the phosphate is still bound to the motor.
  %%     ``Priming stroke?''
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \onebound &
  %%   6 & \begin{minipage}{5in} It is now bound to the microtubule, but
  %%     with low affinity, and without any loss of phosphate. Question:
  %%     does this happen many times? If so it may still be onebound.
  %%     Otherwise it is probably bothbound.
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \bothbound &
  %%   7 & \begin{minipage}{5in}
  %%     Now it is seriously bound to the microtubule, and the motor
  %%     domain has changed conformation.  This is now ``high affinity.''
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \bothbound &
  %%   8 & \begin{minipage}{5in}
  %%     Phosphate has been released, and our equilibrium angles have
  %%     been changed.  This is the ``power stroke.''
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \bothbound &
  %%   1 & \begin{minipage}{5in}
  %%     The ADP has been released from the motor domain, so it can now
  %%     bind ATP.  Still ``high affinity,'' and still in the
  %%     ``power-stroke'' conformation.
  %%   \end{minipage}\vspace{0.5em}\\
  %%   \bothbound &
  %%   9 & \begin{minipage}{5in}
  %%     While stuck in this ``normal bound'' state, another site on the
  %%     motor AAA3 can lose its ADP, which leaves the motor in a state
  %%     where it cannot directly move on to step 2 (until after
  %%     reversing this).  Controversial?
  %%   \end{minipage}
  %% \end{tabular}
  \caption{Cianfrocco mechanochemical cycle.}{Cycle of states the dynein motor goes through during motion in the Cianfrocco model \cite{cianfroccoreview}.}
  \label{mech-cycle}
\end{figure}

%% \subsection{Other models}
%% Find other models and put them here

%% \section{Brownian dynamics}
%% New section which describes how BD works.

\section{Testing the Cianfrocco model}
One of the main goals of studying dynein is establishing how the motor walks. This can be a difficult task, since direct observation of the protein walking is not currently experimentally possible. Given the experimental challenges, a computational approach seems promising. This approach involves forming a hypothesis on how the motor walks, creating a model which behaves according to this hypothesis, simulating that model, and comparing the simulation's behavior to experimental data. This process can be an efficient way to support hypotheses and guide future research.\\

\subsection{Computational approach}
This project seeks to create a dynamic model of dynein and use it to answer questions. The Cianfrocco model provides a rough framework for the events that go on during the dynein mechanochemical cycle. By augmenting this model with information about dynein's structure and conformation changes occuring during the cycle, a new dynamic model which takes into account both temporal and spatial information can be constructed. This dynamic model can then be simulated under cellular conditions to test how it behaves. Such a model would be ideal, since it would make very precise predictions about how real dynein walks. This simulation would act as an experiment testing the validity of the Cianfrocco model.\\

>>>>

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}
% \readytosubmit %% Uncomment this line before submitting, so that the instruction page is removed.

\title{Your Main Manuscript Title}
\author{Author1, Author2 and Author3 (Complete author list)}
\correspondingauthor{Corresponding Author Name.\\E-mail: author.two@email.com}

@@

%% Comment/remove this line before generating final copy for submission
\instructionspage

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\section*{Deriving motion equations}
\subsection*{Onebound case}
The following equations allow recursive calculation of the positions of each domain in Cartesian coordinates:
